---
title: "Project 2"
output: pdf_document
---


# 1. Data Collection and Exploration

```{r}
df1 <- read.table(file="/Users/stellapark95/Desktop/STAT154/projects/project2/image_data/image1.txt")

df2 <- read.table(file="/Users/stellapark95/Desktop/STAT154/projects/project2/image_data/image2.txt")

df3 <- read.table(file="/Users/stellapark95/Desktop/STAT154/projects/project2/image_data/image3.txt")
```



```{r}
df <- rbind(df1, df2, df3)
```

```{r}
colnames(df) <- c('y', 'x', 'expert', 'ndai', 'sd', 'corr', 'df', 'cf', 'bf', 'af', 'an')
colnames(df1) <- c('y', 'x', 'expert', 'ndai', 'sd', 'corr', 'df', 'cf', 'bf', 'af', 'an')
colnames(df2) <- c('y', 'x', 'expert', 'ndai', 'sd', 'corr', 'df', 'cf', 'bf', 'af', 'an')
colnames(df3) <- c('y', 'x', 'expert', 'ndai', 'sd', 'corr', 'df', 'cf', 'bf', 'af', 'an')


```


**1(b)**

```{r}
summary(df)
```


```{r}
# percentage per class
print("All")
dim(subset(df, expert==0))[1] / dim(df)[1]
dim(subset(df, expert==1))[1] / dim(df)[1]
dim(subset(df, expert==-1))[1] / dim(df)[1]
print("1")
dim(subset(df1, expert==0))[1] / dim(df1)[1]
dim(subset(df1, expert==1))[1] / dim(df1)[1]
dim(subset(df1, expert==-1))[1] / dim(df1)[1]
print("2")
dim(subset(df2, expert==0))[1] / dim(df2)[1]
dim(subset(df2, expert==1))[1] / dim(df2)[1]
dim(subset(df2, expert==-1))[1] / dim(df2)[1]
print("3")
dim(subset(df3, expert==0))[1] / dim(df3)[1]
dim(subset(df3, expert==1))[1] / dim(df3)[1]
dim(subset(df3, expert==-1))[1] / dim(df3)[1]

```












```{r}
summary(df)
```


```{r}


library(ggplot2)


```



```{r}
ggplot(df1, aes(x, y)) + geom_point(aes(colour = factor(expert)))
```



```{r}
ggplot(df2, aes(x, y)) + geom_point(aes(colour = factor(expert)))
```





```{r}
ggplot(df3, aes(x, y)) + geom_point(aes(colour = factor(expert)))
```



```{r}
ggplot(df, aes(x, y)) + geom_point(aes(colour = factor(expert)))
```




The dataset cannot be iid because the points are correlated to each other. For instance, if one dataset is classified as "cloudy", the dataset near it would be more likely to be classified as "cloudy" as well. 









**1(c)**

```{r}
pairs(df1[4:6])
```






```{r}
pairs(df2[4:6])
```







```{r}
pairs(df3[4:6])
```




```{r}
pairs(df[4:6])
```











```{r}
library(ggplot2)
df1$expert <- factor(df1$expert)
ggplot(df1, aes(expert, ndai)) + geom_violin() + geom_boxplot(width=0.1) + stat_summary(fun.y=mean, geom="point", shape=23, size=2, color='red') + ggtitle('Image 1 NDAI vs. Expert Label')

```



```{r}

df2$expert <- factor(df2$expert)
ggplot(df2, aes(expert, ndai)) + geom_violin() + geom_boxplot(width=0.1) + stat_summary(fun.y=mean, geom="point", shape=23, size=2, color='red') + ggtitle('Image 2 NDAI vs. Expert Label')

```



```{r}

df3$expert <- factor(df3$expert)
ggplot(df3, aes(expert, ndai)) + geom_violin() + geom_boxplot(width=0.1) + stat_summary(fun.y=mean, geom="point", shape=23, size=2, color='red') + ggtitle('Image 3 NDAI vs. Expert Label')

```






# 2. Preparation


**2(a)**

```{r}
t_df <- subset(x = df, expert==1 | expert==-1)
t_df1 <- subset(x = df1, expert==1 | expert==-1)
t_df2 <- subset(x = df2, expert==1 | expert==-1)
t_df3 <- subset(x = df3, expert==1 | expert==-1)

```

### Split Method 1

```{r}
library(lattice)
library(caret)
```





* Image 1

```{r}
df1_q1 <- subset(x = t_df1, x >= mean(x) & y >= mean(y))
df1_q2 <- subset(x = t_df1, x < mean(x) & y >= mean(y))
df1_q3 <- subset(x = t_df1, x < mean(x) & y < mean(y))
df1_q4 <- subset(x = t_df1, x >= mean(x) & y < mean(y))

```

  * Quadrant 1

```{r}
set.seed(42)
df1_q1_i <- createDataPartition(df1_q1$expert, p=0.8, list=FALSE)
df1_q1_t <- df1_q1[df1_q1_i,]
df1_q1_test <- df1_q1[-df1_q1_i,]

df1_q1_ti <- createDataPartition(df1_q1_train$expert, p=0.75, list=FALSE)
df1_q1_train <- df1_q1_t[df1_q1_ti, ]
df1_q1_val <- df1_q1_t[-df1_q1_ti, ]

```




  * Quadrant 2

```{r}
set.seed(42)
df1_q2_i <- createDataPartition(df1_q2$expert, p=0.8, list=FALSE)
df1_q2_t <- df1_q2[df1_q2_i,]
df1_q2_test <- df1_q2[-df1_q2_i,]


df1_q2_ti <- createDataPartition(df1_q2_train$expert, p=0.75, list=FALSE)
df1_q2_train <- df1_q2_t[df1_q2_ti, ]
df1_q2_val <- df1_q2_t[-df1_q2_ti, ]


```

  * Quadrant 3

```{r}
set.seed(42)
df1_q3_i <- createDataPartition(df1_q3$expert, p=0.8, list=FALSE)
df1_q3_t <- df1_q3[df1_q3_i,]
df1_q3_test <- df1_q3[-df1_q3_i,]

df1_q3_ti <- createDataPartition(df1_q3_train$expert, p=0.75, list=FALSE)
df1_q3_train <- df1_q3_t[df1_q3_ti, ]
df1_q3_val <- df1_q3_t[-df1_q3_ti, ]

```

  * Quadrant 4

```{r}
set.seed(42)
df1_q4_i <- createDataPartition(df1_q4$expert, p=0.8, list=FALSE)
df1_q4_t <- df1_q4[df1_q4_i,]
df1_q4_test <- df1_q4[-df1_q4_i,]

df1_q4_ti <- createDataPartition(df1_q4_train$expert, p=0.75, list=FALSE)
df1_q4_train <- df1_q4_t[df1_q4_ti, ]
df1_q4_val <- df1_q4_t[-df1_q4_ti, ]

```




* Image 2

```{r}
df2_q1 <- subset(x = t_df2, x >= mean(x) & y >= mean(y))
df2_q2 <- subset(x = t_df2, x < mean(x) & y >= mean(y))
df2_q3 <- subset(x = t_df2, x < mean(x) & y < mean(y))
df2_q4 <- subset(x = t_df2, x >= mean(x) & y < mean(y))

```


  * Quadrant 1

```{r}
set.seed(42)
df2_q1_i <- createDataPartition(df2_q1$expert, p=0.8, list=FALSE)
df2_q1_t <- df2_q1[df2_q1_i,]
df2_q1_test <- df2_q1[-df2_q1_i,]

df2_q1_ti <- createDataPartition(df2_q1_train$expert, p=0.75, list=FALSE)
df2_q1_train <- df2_q1_t[df2_q1_ti, ]
df2_q1_val <- df2_q1_t[-df2_q1_ti, ]

```





  * Quadrant 2

```{r}
set.seed(42)
df2_q2_i <- createDataPartition(df2_q2$expert, p=0.8, list=FALSE)
df2_q2_t <- df2_q2[df2_q2_i,]
df2_q2_test <- df2_q2[-df2_q2_i,]

df2_q2_ti <- createDataPartition(df2_q2_train$expert, p=0.75, list=FALSE)
df2_q2_train <- df2_q2_t[df2_q2_ti, ]
df2_q2_val <- df2_q2_t[-df2_q2_ti, ]

```

  * Quadrant 3

```{r}
set.seed(42)
df2_q3_i <- createDataPartition(df2_q3$expert, p=0.8, list=FALSE)
df2_q3_t <- df2_q3[df2_q3_i,]
df2_q3_test <- df2_q3[-df2_q3_i,]

df2_q3_ti <- createDataPartition(df2_q3_train$expert, p=0.75, list=FALSE)
df2_q3_train <- df2_q3_t[df2_q3_ti, ]
df2_q3_val <- df2_q3_t[-df2_q3_ti, ]

```

  * Quadrant 4

```{r}
set.seed(42)
df2_q4_i <- createDataPartition(df2_q4$expert, p=0.8, list=FALSE)
df2_q4_t <- df2_q4[df2_q4_i,]
df2_q4_test <- df2_q4[-df2_q4_i,]

df2_q4_ti <- createDataPartition(df2_q4_train$expert, p=0.75, list=FALSE)
df2_q4_train <- df2_q4_t[df2_q4_ti, ]
df2_q4_val <- df2_q4_t[-df2_q4_ti, ]

```





* Image 3

```{r}
df3_q1 <- subset(x = t_df3, x >= mean(x) & y >= mean(y))
df3_q2 <- subset(x = t_df3, x < mean(x) & y >= mean(y))
df3_q3 <- subset(x = t_df3, x < mean(x) & y < mean(y))
df3_q4 <- subset(x = t_df3, x >= mean(x) & y < mean(y))

```



  * Quadrant 1

```{r}
set.seed(42)
df3_q1_i <- createDataPartition(df3_q1$expert, p=0.8, list=FALSE)
df3_q1_t <- df3_q1[df3_q1_i,]
df3_q1_test <- df3_q1[-df3_q1_i,]

df3_q1_ti <- createDataPartition(df3_q1_train$expert, p=0.75, list=FALSE)
df3_q1_train <- df3_q1_t[df3_q1_ti, ]
df3_q1_val <- df3_q1_t[-df3_q1_ti, ]

```





  * Quadrant 2

```{r}
set.seed(42)
df3_q2_i <- createDataPartition(df3_q2$expert, p=0.8, list=FALSE)
df3_q2_t <- df3_q2[df3_q2_i,]
df3_q2_test <- df3_q2[-df3_q2_i,]

df3_q2_ti <- createDataPartition(df3_q2_train$expert, p=0.75, list=FALSE)
df3_q2_train <- df3_q2_t[df3_q2_ti, ]
df3_q2_val <- df3_q2_t[-df3_q2_ti, ]

```

  * Quadrant 3

```{r}
set.seed(42)
df3_q3_i <- createDataPartition(df3_q3$expert, p=0.8, list=FALSE)
df3_q3_t <- df3_q3[df3_q3_i,]
df3_q3_test <- df3_q3[-df3_q3_i,]

df3_q3_ti <- createDataPartition(df3_q3_train$expert, p=0.75, list=FALSE)
df3_q3_train <- df3_q3_t[df3_q3_ti, ]
df3_q3_val <- df3_q3_t[-df3_q3_ti, ]

```

  * Quadrant 4

```{r}
set.seed(42)
df3_q4_i <- createDataPartition(df3_q4$expert, p=0.8, list=FALSE)
df3_q4_t <- df3_q4[df3_q4_i,]
df3_q4_test <- df3_q4[-df3_q4_i,]

df3_q4_ti <- createDataPartition(df3_q4_train$expert, p=0.75, list=FALSE)
df3_q4_train <- df3_q4_t[df3_q4_ti, ]
df3_q4_val <- df3_q4_t[-df3_q4_ti, ]

```




```{r}
train_1 <- rbind(
  df1_q1_train,
  df1_q2_train,
  df1_q3_train,
  df1_q4_train,
  
  df2_q1_train,
  df2_q2_train,
  df2_q3_train,
  df2_q4_train,
  
  df3_q1_train,
  df3_q2_train,
  df3_q3_train,
  df3_q4_train
)

test_1 <- rbind(
  df1_q1_test,
  df1_q2_test,
  df1_q3_test,
  df1_q4_test,
  
  df2_q1_test,
  df2_q2_test,
  df2_q3_test,
  df2_q4_test,
  
  df3_q1_test,
  df3_q2_test,
  df3_q3_test,
  df3_q4_test
)
  
val_1 <- rbind(
  df1_q1_val,
  df1_q2_val,
  df1_q3_val,
  df1_q4_val,
  
  df2_q1_val,
  df2_q2_val,
  df2_q3_val,
  df2_q4_val,
  
  df3_q1_val,
  df3_q2_val,
  df3_q3_val,
  df3_q4_val
)
  
```















